---
title: "Mental Health Outcomes in Gamers"
author: "Team 5 -- DATS 6101"
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
bibliography: reference.bib
csl: "apa.csl"
---

```{r init, include=F}
library(ezids)
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(rstatix)
library(knitr)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(patchwork)
library(randomForest)
library(smotefamily)
library(caret)
library(ranger)
```


```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
options(scientific=T, digits = 3) 
```

```{r read-in, include=TRUE}
df <- read.csv("GamingStudy_data.csv")
```


## <u>Introduction</u>
Video games are a common pastime for today's generation. With the various advancements in graphics and the internet itself, the immersion of video games continues to increase. One topic we wanted to research was the gaming habits of people with psychological struggles. 

### Reasearch Question

Does a person’s video game habits, seen in hours spent playing and streaming, affect their mental health?

### The Data

This data set (https://www.kaggle.com/datasets/divyansh22/online-gaming-anxiety-data/data) surveyed 13,464 gamers and had them answer questions not only about time spent playing video games, but it also asked questions that are typically seen in psychological diagnosis tests approved in the Diagnostic and Statistical Manual of Mental Disorders (DSM) [@marian_sauter_gaming_2017]. Please note that the terms gamers and surveyors will be used interchangeably in this report.

### Mental Health Inventories and Predictors 

This is the list of mental health indicators.

1. GAD-7 (Generalized Anxiety Disorder, 7 questions)

2. GADE (Generalized Anxiety Disorder, Effect in Life)

3. SWL (Satisfaction with Life)

4. SPIN (Social Phobia Inventories)

This is the list of mental health predictors.

1. Hours: the hours per week playing video games

2. Streams: the hours per week spent streaming video games

3. whyplay: the reason why surveyors play video games

4. Age: the number of years a surveyor is when the survey was completed

5. Gender: the sex of surverors

6. Game Played: the selected game surveyors play the most


## <u>Exploratory Data Analysis</u>

In order to start answering our research question, we need to examine our data. This will include investigating demographic data, psychological inventory scores, and ensuring standardized metrics for times and reasoning.

### Assumption of Continuity

In our EDA, we noticed that the data for the psychological inventories was not truly continuous. This was made apparent when attempting logarithmic and scale transformations. However, due to the level of difficulty this course is expecting us to perservere, we are assuming normality for all variables.

### Demographics

First, we need to examine the demographics of surveyors. Since we have age, gender, country of residence, and game played, let's see what we can find out.

``` {r, age}
age_counts <- as.data.frame(table(df$Age))
colnames(age_counts) <- c("Age", "SurveyCount")
age_counts <- age_counts[order(-age_counts$SurveyCount), ]

# Create the plot
fig <- ggplot(age_counts, aes(x = reorder(Age, -SurveyCount), y = SurveyCount)) +
  geom_bar(stat = 'identity', fill = '#232A75') +
  labs(title = 'Age Distribution', 
       subtitle = '', 
       x = 'Age', 
       y = 'Total Respondents') +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_minimal()  +
 theme(plot.title = element_text(hjust = 0.5))

# Print the figure
print(fig)

```

The age histogram tells us that gamers who are 18 years old make up most of the surveyors, followed by 19 and 20 year old gamers. This means age is extremely right skewed.

``` {r, genderplot}
gender_counts <- as.data.frame(table(df$Gender))
colnames(gender_counts) <- c("Gender", "SurveyCount")
gender_counts <- gender_counts[order(-gender_counts$SurveyCount), ]

# Create the plot
fig <- ggplot(gender_counts, aes(x = reorder(Gender, -SurveyCount), y = SurveyCount)) +
  geom_bar(stat = 'identity', fill = '#232A75') +
  labs(title = 'Gender Distribution', 
       subtitle = '', 
       x = '', 
       y = 'Total Respondents') +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_minimal()  +
 theme(plot.title = element_text(hjust = 0.5))

# Print the figure
print(fig)
```

The respondents/surveyors are almost entirely males. This is another, significant right-skewed variable.

``` {r, countryplot}
Residence_counts <- as.data.frame(table(df$Residence))
colnames(Residence_counts) <- c("Residence", "SurveyCount")
Residence_counts <- Residence_counts[order(-Residence_counts$SurveyCount), ]

# Keep only the top 10 residences
top_10_residences <- head(Residence_counts, 10)

# Create the plot
fig <- ggplot(top_10_residences, aes(x = reorder(Residence, -SurveyCount), y = SurveyCount)) +
  geom_bar(stat = 'identity', fill = '#232A75') +
  geom_text(aes(label = SurveyCount), vjust = -0.5, color = 'white') +
  labs(title = 'Top 10 Residences Distribution', 
       x = '', 
       y = 'Total Surveyors') +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Print the figure
print(fig)
```

Interestingly, we see that most of the surveyors reside in the United States of America. Since there were a lot of countries available for selection in the survey, we limited the histogram to only show the top 10 countries.

``` {r, gameplot}
game_counts <- as.data.frame(table(df$Game))
colnames(game_counts) <- c("Game", "SurveyCount")
game_counts <- game_counts[order(-game_counts$SurveyCount), ]

# Create the plot
fig <- ggplot(game_counts, aes(x = reorder(Game, -SurveyCount), y = SurveyCount)) +
  geom_bar(stat = 'identity', fill = '#232A75') +
  labs(title = 'Game Distribution', 
       subtitle = '', 
       x = '', 
       y = 'Total Surveyors') +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_minimal()  +
 theme(plot.title = element_text(hjust = 0.5))
# Print the figure
print(fig)
```

Finally, we see that the game most played is League of Legends, a multi-player battle arena video game. While this also makes this variable right-skewed, it also informs us that the data analyzed for this project is applicable primarily for gamers who engage in multi-player/online mediums. League of Legends happens to be the best represented in our data.

To summarize the demographics of our data, we are mostly looking at 18 year old males that reside in the United States and play League of Legends.

### Psychological Inventories

Now that we know the demographics of our data, it's time to look at the scores of the four psychological inventories. 

```{r, gad_dist}
ggplot(data = df, aes(x = GAD_T)) +
  geom_histogram(binwidth = 1,
                 color = "red",
                 fill = "blue",
                 alpha = .7) + # opacity
  labs(x = "GAD Total Score", y = "Frequency", title = "GAD Distribution") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

For GAD-7, interestingly, most of the total scores are either zero or two. Scores from zero to four indicate minimal symptoms of anxiety. Scores from five to nine indicate mild symptoms of anxiety, potentially indicating sub-clinical GAD. Scores from ten to fourteen indicate moderate symptoms of anxiety, likely indicating GAD. Scores from fifteen to twenty-one indicate severe symptoms of anxiety, strongly suggesting GAD. This right skew means that there are a lot of surveyors who experience minimal symptoms of generalized anxiety. 

```{r, gade_dist}
# Reorder the GADE factor levels
df$GADE <- factor(df$GADE, levels = c("Not difficult at all", "Somewhat difficult", "Very difficult", "Extremely difficult", NA))

# Plot the histogram of reordered GADE values
ggplot(df, aes(x = GADE)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Histogram of GADE Values",
       x = "GADE (Difficulty Level)",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

For GADE, there are a lot of surveyors that have indicated they can function "appropriately" in society, shown in the 'Not difficult at all' bar. This means that the effect generalized anxiety may have in their life is extremely minimal. The data here is very right-skewed, as with the other variables where EDA has been performed.

```{r, swl_dist}
ggplot(data = df, aes(x = SWL_T)) +
  geom_histogram(binwidth = 1,
                 color = "pink",
                 fill = "green",
                 alpha = .7) + # opacity
  labs(x = "SWL Total Score", y = "Frequency", title = "SWL Distribution") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

For SWL, this appears to be the most normally distributed, with most surveyors having scores in the low to mid twenties. The higher the score, the more satisfied with life a person is. The lower the score, the more dissatisfied with life a person is.

```{r, spin_dist}
ggplot(data = df, aes(x = SPIN_T)) +
  geom_histogram(binwidth = 1,
                 color = "orange",
                 fill = "magenta",
                 alpha = .7) + # opacity
  labs(x = "SPIN Total Score", y = "Frequency", title = "SPIN Distribution") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

For SPIN, the data is slightly less skewed; we see that the most frequent score is around 14 or 15, indicating minimal social anxiety. The lower the score, the less likely one experiences social anxieties and phobias. The higher the score, the more likely one experiences social anxieties and phobias.

### Standarized Metrics for Psychological Inventories

Immediately we see that the total scores for all the psychological inventories need to be standardized into their respective categories of minimal and severe. Once the conversion of raw scores into ordinal values is complete, we can get a better representation of how each psychological inventory affects gamers. The conversion is based on some research articles [@greenspace_mental_health_social_2020; @yun_satisfaction_2019; @sapra_using_2020].

``` {r, ord}
# Function to convert GAD_T to ordinal variable
convert_GAD <- function(score) {
  if (is.na(score)) return(NA)  # Handle NA values
  if (score >= 0 & score <= 4) {
    return("minimal anxiety")
  } else if (score >= 5 & score <= 9) {
    return("mild anxiety")
  } else if (score >= 10 & score <= 14) {
    return("moderate anxiety")
  } else if (score >= 15 & score <= 21) {
    return("severe anxiety")
  } else {
    return(NA)  # Handles scores outside defined range
  }
}

# Function to convert SWL_T to ordinal variable
convert_SWL <- function(score) {
  if (is.na(score)) return(NA)  # Handle NA values
  if (score >= 31 & score <= 35) {
    return("extremely satisfied")
  } else if (score >= 26 & score <= 30) {
    return("satisfied")
  } else if (score >= 21 & score <= 25) {
    return("slightly satisfied")
  } else if (score == 20) {
    return("neutral")
  } else if (score >= 15 & score <= 19) {
    return("slightly dissatisfied")
  } else if (score >= 10 & score <= 14) {
    return("dissatisfied")
  } else if (score >= 5 & score <= 9) {
    return("extremely dissatisfied")
  } else {
    return(NA)  # Handles scores outside defined range
  }
}

# Function to convert SPIN_T to ordinal variable
convert_SPIN <- function(score) {
  if (is.na(score)) return(NA)  # Handle NA values
  if (score >= 0 & score <= 20) {
    return("none")
  } else if (score >= 21 & score <= 30) {
    return("mild")
  } else if (score >= 31 & score <= 40) {
    return("moderate")
  } else if (score >= 41 & score <= 50) {
    return("severe")
  } else if (score >= 51 & score <= 68) {
    return("very severe")
  } else {
    return(NA)  # Handles scores outside defined range
  }
}

# Function to convert hours to ordinal variable
convert_hours <- function(hours) {
  if (is.na(hours)) return(NA)  # Handle NA values
  if (hours >= 0 & hours <= 10) {
    return("Low")
  } else if (hours > 10 & hours <= 20) {
    return("Moderate")
  } else if (hours > 20 & hours <= 30) {
    return("High")
  } else if (hours > 30) {
    return("Very High")
  } else {
    return(NA)  # Handles any unexpected values
  }
}

# Create new columns in the data frame
df$GAD_ord <- sapply(df$GAD_T, convert_GAD)
df$SWL_ord <- sapply(df$SWL_T, convert_SWL)
df$SPIN_ord <- sapply(df$SPIN_T, convert_SPIN)
df$hours_ord <- sapply(df$Hours, convert_hours)


#make values ordinal (min to max)
df$GAD_ord <- factor(df$GAD_ord, 
                     levels = c("minimal anxiety", "mild anxiety", "moderate anxiety", "severe anxiety"), 
                     ordered = TRUE)

df$SWL_ord <- factor(df$SWL_ord, 
                     levels = c("extremely satisfied", "satisfied", "slightly satisfied", "neutral", 
                                "slightly dissatisfied", "dissatisfied", "extremely dissatisfied"), 
                     ordered = TRUE)

df$SPIN_ord <- factor(df$SPIN_ord, 
                      levels = c("none", "mild", "moderate", "severe", "very severe"), 
                      ordered = TRUE)

df$hours_ord <- factor(df$hours_ord, 
                            levels = c("Low", "Moderate", "High", "Very High"), 
                            ordered = TRUE)


# Print the updated data frame
#print(df$GAD_ord)
```

```{r GAD_ord_dist}
ggplot(df, aes(x = GAD_ord)) +
  geom_bar(fill = "blue") +
  labs(title = "Histogram of Ordinal GAD Values",
       x = "Ordinal GAD Values",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

GAD is still right-skewed, but we can see that most gamers experience minimal anxiety. Since GADE has already been displayed in the correct order, we can move onto SWL.

```{r SWL_ord_dist}
ggplot(df, aes(x = SWL_ord)) +
  geom_bar(fill = "green") +
  labs(title = "Histogram of Ordinal SWL Values",
       x = "Ordinal SWL Values",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

SPIN is interestingly bi-modal, with 'Slightly satisfied' and 'Slightly dissatisfied' being the most common selections chosen.

```{r SPIN_ord_dist}
ggplot(df, aes(x = SPIN_ord)) +
  geom_bar(fill = "magenta") +
  labs(title = "Histogram of Ordinal SPIN Values",
       x = "Ordinal SPIN Values",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

SPIN is right-skewed, where most gamers indicated they experience no social anxieties or phobias.

### Reason for playing

When looking at all of this data, a crucial component is the reason why gamers play video games in the first place. Since this field was a free-response, it would be impossible, and unwise, to display a histogram of every possible response received. In order to show meaningful results, we will use regular expressions to isolate key words in the responses and categorize them under a specific group.

``` {r, cleaningwhyplay}
clean_responses <- function(response) {
  response <- trimws(tolower(response))  # Normalize to lowercase and trim whitespace
  
  # Check for exact matches first
  if (response %in% c("winning", "improving", "relaxing", "having fun")) {
    return(response)
  }
  
  # Check for various conditions to set the response
  if (str_detect(response, "^a[\\s,\\.]", ) || 
      str_detect(response, "^1[\\s,\\.]", ) || 
      str_detect(response, "^all") || 
      str_detect(response, "^mix") || 
      str_detect(response, "combination") || 
      str_detect(response, "everything")) {
    return("all of the above")
  }
  
  if (str_detect(response, "distraction|escaping|escape|forget|hide|hiding|depression|boredom|bored|time|forgetting|making my brain go numb|distracting|doing something|getting away from real life|trying to get my mind off")) {
    return("escape")
  }

  if (str_detect(response, "improve|improving|learn|learning")) {
    return("improving")
  }
  
  if (str_detect(response, "well|perform|performing")) {
    return("winning")
  }
  
  if (str_detect(response, "social|socializing|friends|team|cooperating|people|the people i play with, the game is less important")) {
    return("social")
  }
  
  if (str_detect(response, "win|winning|beating|victory|competeing|competing|competition|conpeting")) {
    return("winning")
  }
  
  if (str_detect(response, "winning") && str_detect(response, "improving")) {
    return("mix")
  }
  
  if (str_detect(response, "relaxing") && str_detect(response, "fun")) {
    return("mix")
  }
  
  if (str_detect(response, "combination|all")) {
    return("mix")
  }
  
  if (str_detect(response, "fun")) {
    return("having fun")
  }
  
  # Default to original response if no condition is met
  return("bad response")
}

# Apply the cleaning function to the 'whyplay' column
df$cleaned_whyplay <- sapply(df$whyplay, clean_responses)

# Count the occurrences of each cleaned response
whyplay_counts <- as.data.frame(table(df$cleaned_whyplay))
```

```{r, why }

colnames(whyplay_counts) <- c("WhyPlay", "SurveyCount")


whyplay_counts <- whyplay_counts[order(-whyplay_counts$SurveyCount), ]


whyplay_counts$WhyPlay <- factor(whyplay_counts$WhyPlay, levels = whyplay_counts$WhyPlay)


fig <- ggplot(whyplay_counts, aes(x = WhyPlay, y = SurveyCount)) +
  geom_bar(stat = 'identity', fill = '#232A75') +
  labs(title = 'Reasons for Playing', 
       subtitle = '', 
       caption = 'Data Source: Survey Data', 
       x = 'Reasons for Playing', 
       y = 'Total Respondents') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))


print(fig)
```

Now we can see that most gamers play either for fun or to improve their skills. One quick note: the group 'Escape' does not refer to escaping responsibilities. It is more along the lines of dissociation and depression.

### Hours Played

Now that we know why gamers play video games, we want to see how long they play per week. However, we want to first see the data shown in categories of low (0 - 10 hours), moderate (11 - 20 hours), high (21 -30 hours), and very high (over 30 hours). 

```{r hours_ord_dist}
ggplot(df, aes(x = hours_ord)) +
  geom_bar(fill = "turquoise") +
  labs(title = "Histogram of Ordinal Hours Playing Video Games per Week",
       x = "Ordinal Hours Values",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

We can see that most gamers play from 11 - 20 hours per week, which is the moderate group. This distribution is slightly less skewed compared to the other variables where EDA has been performed. 

### Identifying outliers in Hours

In order to find relationship between time spent playing games and its effect on a person's mental health, we need to explore the variable Hours spent (per week) playing video games. We generate a histogram to see its distribution.

```{r}

# Replace missing values in the Hours column with NA explicitly
hrs_data_cleaned <- df %>%
  mutate(Hours = ifelse(is.na(Hours), NA, Hours))
missing_values <- colSums(is.na(hrs_data_cleaned))

# Filter necessary columns
hrs_GAD_filtered <- hrs_data_cleaned %>%
  select(Hours, GAD1:GAD7, GAD_T)

# Finding outliers in Hours
fig_hours <- ggplot(hrs_GAD_filtered, aes(x = Hours)) + 
  geom_boxplot(colour = "orange", fill = "#7777cc", outlier.colour = "red", outlier.shape = 8, outlier.size = 5) +
  labs(title = "Outliers in Hours spent gaming per week", 
       subtitle = '', 
       caption = 'Data Source: Survey Data', 
       x = "Hours Spent (per week)", 
       y = '') +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center and bold title
    axis.title.x = element_text(hjust = 0.5), # Bold and center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Bold and center y-axis label
  )

print(fig_hours)

```

We see that the data sure has anomaly because it's impossible to have 8000 hours in a week! So in order to use this variable, such extreme values need to be removed.

### Removing outliers in Hours and generating Q-Q plot, histogram and boxplot

We now use a function to remove the extreme values from Hours played variable and generate QQ-plot, histogram and boxplot and check the data distribution again.

```{r}
# Define the outlierKD2 function
outlierKD2 <- function(dt, var, rm = TRUE, boxplt = TRUE, qqplt = TRUE) { 
  var_name <- eval(substitute(var),eval(dt))
  na1 <- sum(is.na(var_name))
  m1 <- mean(var_name, na.rm = TRUE)
  
  # Generate boxplot and histogram (before removing outliers)
  if (boxplt) {
    par(mfrow = c(2, 2), oma = c(0, 0, 3, 0))
    boxplot(var_name, main = "With outliers")
    hist(var_name, main = "With outliers", xlab = NA, ylab = NA)
  }
  
  outlier <- boxplot.stats(var_name)$out
  mo <- mean(outlier)
  
  var_name <- ifelse(var_name %in% outlier, NA, var_name)
  
  # Generate boxplot and histogram (after removing outliers)
  if (boxplt) {
    boxplot(var_name, main = "Without outliers")
    hist(var_name, main = "Without outliers", xlab = NA, ylab = NA)
  }
  
  # Generate QQ plot if requested
  if (qqplt) {
    qqnorm(var_name)
    qqline(var_name)
  }
  
  na2 <- sum(is.na(var_name))
  
  cat("Outliers identified:", na2 - na1, "\n")
  cat("Proportion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name)) * 100, 1), "\n")
  cat("Mean of the outliers:", round(mo, 2), "\n")
  m2 <- mean(var_name, na.rm = TRUE)
  cat("Mean without removing outliers:", round(m1, 2), "\n")
  cat("Mean if we remove outliers:", round(m2, 2), "\n")
  
  # Option to remove outliers (replace with NA)
  if (rm) {
    dt[[as.character(substitute(var))]] <- invisible(var_name)
    assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
    cat("Outliers successfully removed\n")
    return(invisible(dt))
  } else {
    cat("Nothing changed\n")
    return(invisible(var_name))
  }
}

# Apply outlierKD2 to remove outliers from 'Hours'
hrs_GAD_filtered_clean <- outlierKD2(hrs_GAD_filtered, Hours, rm = TRUE, boxplt = FALSE, qqplt = TRUE)

# Create a histogram using ggplot
ggplot(hrs_GAD_filtered_clean, aes(x = Hours)) + 
  geom_histogram(fill = "lightblue", binwidth = 1, color = "black") +
  labs(title = "Histogram of Hours Spent Gaming (Outliers Excluded)", 
       x = "Hours Spent (per week)", 
       y = "Frequency", 
       caption = "Data Source: Survey Data") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center and bold title
    axis.title.x = element_text(hjust = 0.5), # Bold and center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Bold and center y-axis label
  )

# Create a boxplot using ggplot
ggplot(hrs_GAD_filtered_clean, aes(x = Hours)) + 
  geom_boxplot(colour = "orange", fill = "#7777cc", outlier.colour = "red", outlier.shape = 8, outlier.size = 4) +
  labs(title = "Hours spent Gaming per week (Outliers Excluded)", 
       x = "Hours Spent (per week)", 
       y = "", 
       caption = "Data Source: Survey Data") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center and bold title
    axis.title.x = element_text(hjust = 0.5), # Bold and center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Bold and center y-axis label
  )


```

After cleaning the Hours played variable, the Q-Q plot suggests that the center of the data could be more or less normally distributed, but the tails have more or fewer outliers. The histogram confirms that as we can see most frequent values are in the middle portion of the data. The wide shape of the boxplot confirms that the data is widely spread out and the median being close to the upper end indicates that the data is slightly left-skewed.  


### Streams
### Identifying outliers in Streams

We also need to explore the variable streams to know the time spent on watching Streams, Tournaments, etc. (per week) which could have relation on a person's health. We generate a histogram to see its distribution.

```{r}
ggplot(df, aes(x = streams)) + 
  geom_boxplot(colour = "orange", fill = "#7777cc", outlier.colour = "red", outlier.shape = 8, outlier.size = 4) +
  labs(title = "Outliers in time spent on streaming (per week)", 
       x = "Time spent on Streaming (per week)", 
       y = "") +
  theme_minimal() +  # Apply minimal theme
  theme(
    plot.title = element_text(hjust = 0.5),  # Center title
    axis.title.x = element_text(hjust = 0.5), # Center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Center y-axis label
  )

```

It is evident from the plot above that similar to the data of time spent on playing online games, the time spent on watching streams per week also has anomalies. We proceed to remove these extreme values in the next step.


### Removing outliers in Streams and generating Q-Q plot, histogram and boxplot

```{r}

# Define the outlierKD2 function
outlierKD2 <- function(dt, var, rm = TRUE, boxplt = TRUE, qqplt = TRUE) { 
  var_name <- eval(substitute(var), eval(dt))
  na1 <- sum(is.na(var_name))
  m1 <- mean(var_name, na.rm = TRUE)
  
  # Generate plots (boxplot and histogram)
  if (boxplt) {
    par(mfrow = c(2, 2), oma = c(0, 0, 3, 0))
    boxplot(var_name, main = "With outliers")
    hist(var_name, main = "With outliers", xlab = NA, ylab = NA)
  }
  
  outlier <- boxplot.stats(var_name)$out
  mo <- mean(outlier)
  
  var_name <- ifelse(var_name %in% outlier, NA, var_name)
  
  if (boxplt) {
    boxplot(var_name, main = "Without outliers")
    hist(var_name, main = "Without outliers", xlab = NA, ylab = NA)
  }
  
  # Generate QQ plot if requested
  if (qqplt) {
    qqnorm(var_name)
    qqline(var_name)
  }
  
  na2 <- sum(is.na(var_name))
  
  cat("Outliers identified:", na2 - na1, "\n")
  cat("Proportion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name)) * 100, 1), "\n")
  cat("Mean of the outliers:", round(mo, 2), "\n")
  m2 <- mean(var_name, na.rm = TRUE)
  cat("Mean without removing outliers:", round(m1, 2), "\n")
  cat("Mean if we remove outliers:", round(m2, 2), "\n")
  
  # Option to remove outliers (replace with NA)
  if (rm) {
    dt[[as.character(substitute(var))]] <- invisible(var_name)
    assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
    cat("Outliers successfully removed\n")
    return(invisible(dt))
  } else {
    cat("Nothing changed\n")
    return(invisible(var_name))
  }
}

# Use outlierKD2 to remove outliers from 'streams'
df_clean <- outlierKD2(df, streams, rm = TRUE, boxplt = FALSE, qqplt = TRUE)

# Histogram of cleaned data
hist(df_clean$streams, main = "Histogram of Time Spent Watching Streams (hrs/week)", 
     xlab = "Time Spent Watching Streams (hrs/week)", col = "lightblue")

# Boxplot of cleaned data
ggplot(df_clean, aes(x = streams)) + 
  geom_boxplot(colour = "orange", fill = "#7777cc", outlier.colour = "red", outlier.shape = 8, outlier.size = 4) +
  labs(title = "Time Spent Streaming per week (Outliers Excluded)", 
       x = "Streaming (per week)", 
       y = "") +
  theme_minimal() +  # Apply minimal theme
  theme(
    plot.title = element_text(hjust = 0.5),  # Center title
    axis.title.x = element_text(hjust = 0.5), # Center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Center y-axis label
  )


```

Even after removing the outliers, we find the data to be not normally distributed (as seen in the Q-Q plot). The histogram confirms this as we find more data towards the lower tail compared to the upper tail. The long length of the whisker of the boxplot confirms this as well.


### Hours and Streams (Pearson's Correlation)

We now conduct the Pearson's correlation test to measure the strength and direction of the linear relationship between Hours spent gaming and Hours spent watching streams:


```{r, corr_h_s}
cor_test <- cor.test(hrs_GAD_filtered_clean$Hours, df_clean$streams)
correlation_result_Hours_Streams <- cor(hrs_GAD_filtered_clean$Hours, df_clean$streams, method = "pearson", use = "complete.obs")
cor_test
```

The Correlation coefficient value (0.0844) indicates a weak positive linear relationship between the two variables, Hours and Streams. Since the correlation is close to 0, the relationship is not strong, meaning that as Hours (time spent gaming) increases, the time spent streaming (or vice versa) increases slightly, but not in a very consistent or strong way.

### Standardizing Hours and Streams cleaned data

Now since the data of both the time spent on playing games and watching streams and tournaments are not normally distributed, we try to standardize these data in order to proceed with the tests that assume normality.


``` {r}
df_clean$streams_Stand <- scale(df_clean$streams)
hrs_GAD_filtered_clean$hours_Stand <- scale(hrs_GAD_filtered_clean$Hours)

```

### Hours spent gaming Q-Q plot and histogram with Standardized values

We generate Q-Q plot and histogram with the Standardized values expecting to get a normal distribution:

```{r, plotting_hours}

# Q-Q plot for standardized Hours
qqnorm(hrs_GAD_filtered_clean$hours_Stand, main = "Q-Q plot for Standardized Hours played gaming")
qqline(hrs_GAD_filtered_clean$hours_Stand)

# Histogram for standardized Hours
ggplot(data = hrs_GAD_filtered_clean, aes(x = hours_Stand)) +
  geom_histogram(binwidth = 1, color = "red", fill = "blue", alpha = .7) +
  labs(x = "Hours played gaming (Standardized)", y = "Frequency", title = "Standardized Hours played gaming distribution (Outliers excluded") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

As seen from the plots above the data now look near to normal and smoother. Now, we look into the time spent watching streams variable. 

### Streaming Q-Q plot and histogram with Standardized values

We generate Q-Q plot and histogram again expecting to get a bell-shaped histogram from the standardized values of the cleaned stream dataset:

``` {r, plotting_streams}

# Q-Q plot for standardized Streams
qqnorm(df_clean$streams_Stand, main = "Q-Q plot for Standardized time spent streaming")
qqline(df_clean$streams_Stand)

# Histogram for standardized Streams
ggplot(data = df_clean, aes(x = streams_Stand)) +
  geom_histogram(binwidth = 1, color = "red", fill = "blue", alpha = .7) +
  labs(x = "Streams Standardized", y = "Frequency", title = "Time spent streaming (Standardized)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

Even after standardizing the streaming time values, the plots depict that the standardized values are not yet normal. Both the plots confirm that the data is more concentrated towards the left side.

### Adding Hours and Streams -> Total Time, and creating histogram and QQ-plot

Previously, We have seen from the Pearson correlation test that there exists a very weak positive correlation when we tried to find a relation between hours and streams variables (which are free of outliers) and the both the variables not normally distributed (even after standardization). So now we tried to combine the two time variables and try to see if the combined data can be used for further analysis.

```{r}

# Create the new column hrs_stream_combined in df
df$hrs_stream_combined <- df_clean$streams + hrs_GAD_filtered_clean$Hours

# Q-Q plot for Total Time
qqnorm(df$hrs_stream_combined, main = "Q-Q plot for Total time")
qqline(df$hrs_stream_combined)

# Histogram with minimal theme and bin width of 5
ggplot(data = df, aes(x = hrs_stream_combined)) +
  geom_histogram(binwidth = 5, color = "red", fill = "lightblue", alpha = 0.7) +
  labs(title = "Histogram: Total Time Spent Watching Streams, Tournaments, etc.", 
       x = "Time Spent (hrs/week)", 
       y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5,),  # Center and bold title
    axis.title.x = element_text(hjust = 0.5), # Bold and center x-axis label
    axis.title.y = element_text(hjust = 0.5)  # Bold and center y-axis label
  )


```

We can clearly see from the generated Q-Q plot and histogram that the combined data called (now called Total time spent on watching streaming and playing games) has now look near to normal. This data is now suitable to conduct all kinds of tests that require normality of data.

### Histogram of Total Time distribution on GAD_ordinal variables

We now try to see the distribution of total time spent on playing games and streaming based on GAD ordinal classification (minimal/ mild/ moderate/ severe levels of anxiety). 

```{r}
ggplot(df, aes(x = df$hrs_stream_combined)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~ GAD_ord) +
  labs(title = "Histogram of Total Time spent on playing games and
       streaming based on GAD ordinal classification",
       x = "Hours Played/Week",
       y = "Count") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center and bold title
    axis.title.x = element_text(hjust = 0.5),  # Center x-axis label
    axis.title.y = element_text(hjust = 0.5)   # Center y-axis label
  )

```

We can see from the above distributions that the hours spent playing games and watching streams is different for different levels of anxiety. This brings us the question if this difference is significantly different or not. We will look into this when doing hypothesis tests.

## <u>Hypothesis Tests</u> 

For hours spent playing and hours spent streaming, an ANOVA test will be conducted to see if there is a significant difference among the different ordinal categories of hours played and hours spent streaming.

For each of the psychological inventories and hours spent playing, ANOVA tests will be conducted to see if there are significant differences among the different ordinal categories of each psychological inventory and hours spent playing.

For each of the psychological inventories, along with GADE, and the reason for playing, chi-squared tests will be conducted to see if the variables are independent of each other.

### Hours Playing and Streaming (ANOVA)

Null Hypothesis (H₀): There is no significant difference in time spent streaming video games and the different hours played categories. $\mu$<sub>low</sub> = $\mu$<sub>moderate</sub> = $\mu$<sub>high</sub> = $\mu$<sub>very high</sub>.

Alternate Hypothesis (H₁): There is a significant difference in  time spent streaming video games and the different hours played categories.

```{r hours_streams_ANOVA}
# Perform one-way ANOVA
anova_result_Hours_Streams <- aov(streams ~ hours_ord, data = df)
p_Hours_Streams <- summary(anova_result_Hours_Streams)[[1]]$`Pr(>F)`[1]
summary(anova_result_Hours_Streams)
```

Since p << 0.05 (p = `r format(p_Hours_Streams, digits = 3)`), we reject the null hypothesis. This means that there is a significant effect hours per week of play time has on a person's total streaming time. In order to see what the effect is, a Tukey's Honest Significant Difference (HSD) post-hoc test must be performed.

```{r TukeyHSD_Hours_Streams}
# Performing Tukey HSD test
tukey_result_Hours_Streams <- TukeyHSD(anova_result_Hours_Streams)

# Print the Tukey HSD results
print(tukey_result_Hours_Streams)

# Plot the Tukey HSD results
plot(tukey_result_Hours_Streams)
```

After performing the post-hoc test, we see that actually there is no significant difference in hours spent streaming and hours spent playing, across all categories.

### GAD and Hours Played (ANOVA)

Null Hypothesis (H₀): There is no significant difference in generalized anxiety disorder scores (GAD_T) across the different hours played categories. $\mu$<sub>low</sub> = $\mu$<sub>moderate</sub> = $\mu$<sub>high</sub> = $\mu$<sub>very high</sub>.

Alternate Hypothesis (H₁): There is a significant difference in generalized anxiety disorder scores across the different hours played categories.

```{r gad_ANOVA}
# Perform one-way ANOVA
anova_result_GAD_Hours <- aov(GAD_T ~ hours_ord, data = df)
p_GAD <- summary(anova_result_GAD_Hours)[[1]]$`Pr(>F)`[1]
summary(anova_result_GAD_Hours)
```

Since p << 0.05 (p = `r format(p_GAD, digits = 3)`), we reject the null hypothesis. This means that there is a significant effect hours per week of play time has on a person's total generalized anxiety disorder score. In order to see what the effect is, a Tukey's Honest Significant Difference (HSD) post-hoc test must be performed.

```{r TukeyHSD_GAD_Hours}
# Performing Tukey HSD test
tukey_result_GAD_Hours <- TukeyHSD(anova_result_GAD_Hours)

# Print the Tukey HSD results
print(tukey_result_GAD_Hours)

# Plot the Tukey HSD results
plot(tukey_result_GAD_Hours)
```

After performing the post-hoc test, we see that there is a significant difference in three categories: Moderate-Low, High-Low, and High-Moderate. We can thereby conclude that there truly is no significant difference in Very High-Low, Very High-Moderate, and Very High-High.

### SWL and Hours Played (ANOVA)

Null Hypothesis (H₀): There is no significant difference in life satisfaction levels (SWL_T) across the different hours played categories. $\mu$<sub>low</sub> = $\mu$<sub>moderate</sub> = $\mu$<sub>high</sub> = $\mu$<sub>very high</sub>.

Alternate Hypothesis (H₁): There is a significant difference in life satisfaction levels  across the different hours played categories.

```{r, swl_ANOVA}
# Perform one-way ANOVA
anova_result_SWL_Hours <- aov(SWL_T ~ hours_ord, data = df)
p_SWL <- summary(anova_result_SWL_Hours)[[1]]$`Pr(>F)`[1]
summary(anova_result_SWL_Hours)
```

Since p << 0.05 (p = `r format(p_SWL, digits = 3)`), we reject the null hypothesis. This means that there is a significant effect hours per week of play time has on a person's total satisfaction with life score. In order to see what the effect is, a Tukey's Honest Significant Difference (HSD) post-hoc test must be performed.

```{r TukeyHSD_SWL_Hours}
# Performing Tukey HSD test
tukey_result_SWL_Hours <- TukeyHSD(anova_result_SWL_Hours)

# Print the Tukey HSD results
print(tukey_result_SWL_Hours)

# Plot the Tukey HSD results
plot(tukey_result_SWL_Hours)
```

After performing the post-hoc test, we see that there is a significant difference in the Moderate-Low category of hours played. We can thereby conclude that there truly is no significant difference in the other five categories. 

### SPIN and Hours Played (ANOVA)

Null Hypothesis (H₀): There is no significant difference in social phobia inventory scores (SPIN_T) across the different hours played categories. $\mu$<sub>low</sub> = $\mu$<sub>moderate</sub> = $\mu$<sub>high</sub> = $\mu$<sub>very high</sub>.

Alternate Hypothesis (H₁): There is a significant difference in social phobia inventory scores across the different hours played categories.

```{r spin_ANOVA}
# Perform one-way ANOVA
anova_result_SPIN_Hours <- aov(SPIN_T ~ hours_ord, data = df)
p_SPIN <- summary(anova_result_SPIN_Hours)[[1]]$`Pr(>F)`[1]
summary(anova_result_SPIN_Hours)
```

Since p << 0.05 (p = `r format(p_SPIN, digits = 3)`), we reject the null hypothesis. This means that there is a significant effect hours per week of play time has on a person's total social phobia inventory score.

```{r TukeyHSD_SPIN_Hours}
# Performing Tukey HSD test
tukey_result_SPIN_Hours <- TukeyHSD(anova_result_SPIN_Hours)

# Print the Tukey HSD results
print(tukey_result_SPIN_Hours)

# Plot the Tukey HSD results
plot(tukey_result_SPIN_Hours)
```

After performing the post-hoc test, we see that there is a significant difference in the Moderate-Low category of hours played. We can thereby conclude that there truly is no significant difference in the other five categories. 

### Inventories and WhyPlay (chi-sqauared)

Based on our theory that the reason why respondents play video games could be an indicators of their mental health, we conducted a chi-squared test of independence against "whyplay" and the inventories. This test determined, with statistical significance, that the "wayplay" and respondents mental health scores on each of the four inventories are dependent (associated).

```{r contingency_GAD, include=TRUE}
contingency_table_gad <- table(df$GAD_ord, df$cleaned_whyplay)


chi_squared_test_gad <- chisq.test(contingency_table_gad)

# Print the results of the Chi-squared test
chi_squared_test_gad
```

```{r contingency_GADE_KEEP, include=TRUE}
# Define the contingency table
contingency_table_GADE <- table(df$GADE, df$cleaned_whyplay)

# Perform the chi-squared test
chi_squared_test_GADE <- chisq.test(contingency_table_GADE)
p_GADE_whyplay <- chi_squared_test_GADE$p.value

# Print the result
chi_squared_test_GADE
```

```{r contingency_SWL, include=TRUE}
contingency_table_swl <- table(df$SWL_ord, df$cleaned_whyplay)


chi_squared_test_swl <- chisq.test(contingency_table_swl)

# Print the results of the Chi-squared test
chi_squared_test_swl
```

```{r contingency_SPIN, include=TRUE}
contingency_table_spin <- table(df$SPIN_ord, df$cleaned_whyplay)



chi_squared_test_spin <- chisq.test(contingency_table_spin)

# Print the results of the Chi-squared test
chi_squared_test_spin
```

In the above chi-squared outputs, we see p-values less than 2.0E-16. These values are several orders of magnitude smaller than 0.05, which tells us that "whyplay" and all three inventories, along with GADE, have some level of dependency.

The following post-hoc analysis calculates the adjusted residuals, which in effect are the difference between the expected and actual output.

#### GAD vs "whyplay"

``` {r, posthoc_GAD}
expected_counts <- chisq.test(contingency_table_gad)$expected

# Calculate residuals
residuals <- contingency_table_gad - expected_counts

# Calculate adjusted residuals
adjusted_residuals <- residuals / sqrt(expected_counts)

# Print adjusted residuals
#print(adjusted_residuals)


residuals_df <- melt(adjusted_residuals)

# Create a heatmap
ggplot(residuals_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "GAD and WhyPlay Adjusted Residuals Heatmap", x = "GAD Scale", y = "WhyPlay") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
```

The GAD vs "whyplay" heatmap clearly shows  that "having fun" is most closely associated "minimal anxiety", while "winning" and "escape" are more associated "moderate" and "severe" anxiety.

```{r GAD_ord vs why_play2}


ggplot(df, aes(x = GAD_ord, fill = cleaned_whyplay)) +
  geom_bar(position = "fill") +  # Use position = "fill" to show proportions
  ylab("Proportion") +           # Label the y-axis
  xlab("Severity of the Generalized Anxiety Disorder") +           # Label the x-axis
  labs(title = "Proportion of Reasons for Playing by GAD Scores",
       fill = "Reasons for Playing") +  # Rename the legend title
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),  # Center and bold the title
    axis.title.x = element_text(face = "bold", size = 12),  # Bold and size for x-axis label
    axis.title.y = element_text(face = "bold", size = 12),  # Bold and size for y-axis label
    axis.text.x = element_text(face = "bold", size = 10),   # Bold the x-axis tick labels
    axis.text.y = element_text(face = "bold", size = 10),   # Bold the y-axis tick labels
    legend.title = element_text(face = "bold"),             # Bold the legend title
    legend.text = element_text(face = "bold")               # Bold the legend labels
  )
```

The above proportions graph offers another way of visualizing the relationship between "whyplay" and GAD. The bar height reflects the proportion of responses within each GAD severity category. This method of visualizations helps us understand the impact of "winning" on the GAD score.

#### GADE vs "whyplay"
 
``` {r, posthoc_GADE}
expected_counts <- chisq.test(contingency_table_GADE)$expected

# Calculate residuals
residuals <- contingency_table_GADE - expected_counts

# Calculate adjusted residuals
adjusted_residuals <- residuals / sqrt(expected_counts)

# Print adjusted residuals
#print(adjusted_residuals)


residuals_df <- melt(adjusted_residuals)

# Create a heatmap
ggplot(residuals_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "GADE and WhyPlay Adjusted Residuals Heatmap", x = "GADE Scale", y = "WhyPlay") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
```

This heatmap showing the residuals for GADE and "whyplay" tell a similar story to that of GAD. "having fun" is highly associated to positive mental health, in the context of anxiety disorders. "winning" and "escaping" as declared reasons why the respondents play video games once again are associated with negative outcomes. Though, "escaping"s affect is considerable higher.

```{r}
ggplot(df, aes(x = GADE, fill = cleaned_whyplay)) +
  geom_bar(position = "fill") +  # Use position = "fill" to show proportions
  ylab("Proportion") +           # Label the y-axis
  xlab("Generalized Anxiety Disorder Effect in Life") +           # Label the x-axis
  labs(title = "Proportion of Reasons for Playing by GADE Scores",
       fill = "Reasons for Playing") +  # Rename the legend title
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),  # Center and bold the title
    axis.title.x = element_text(face = "bold", size = 12),  # Bold and size for x-axis label
    axis.title.y = element_text(face = "bold", size = 12),  # Bold and size for y-axis label
    axis.text.x = element_text(face = "bold", size = 10, angle = 45, hjust = 1),   # Rotate and bold the x-axis tick labels
    axis.text.y = element_text(face = "bold", size = 10),   # Bold the y-axis tick labels
    legend.title = element_text(face = "bold"),             # Bold the legend title
    legend.text = element_text(face = "bold")               # Bold the legend labels
  )


```

We can examine this phenomenon more naively when looking at the proportions graph between GADE and "whyplay". Once again, the relationship of "having fun", "winning", and "escape" to positive and negative outcomes can be seen. 

#### SWL vs "whyplay"

``` {r, posthoc_SWL}
expected_counts <- chisq.test(contingency_table_swl)$expected

# Calculate residuals
residuals <- contingency_table_swl - expected_counts

# Calculate adjusted residuals
adjusted_residuals <- residuals / sqrt(expected_counts)

# Print adjusted residuals
#print(adjusted_residuals)



residuals_df <- melt(adjusted_residuals)

# Create a heatmap
ggplot(residuals_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "SWL and WhyPlay Adjusted Residuals Heatmap", x = "SWL Scale", y = "WhyPlay") +
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

The heatmap above for SWL vs "whyplay" shows that, again, "having fun" is most strongly associated with positive outcomes, satisfaction with life in this case. The response "escape" and "winning" are highly associated with respondents who are extreme dissatisfied in life. 

#### SPIN vs "whyplay"

``` {r, posthoc_SPIN}
expected_counts <- chisq.test(contingency_table_spin)$expected

# Calculate residuals
residuals <- contingency_table_spin - expected_counts

# Calculate adjusted residuals
adjusted_residuals <- residuals / sqrt(expected_counts)

# Print adjusted residuals
#print(adjusted_residuals)


residuals_df <- melt(adjusted_residuals)

# Create a heatmap
ggplot(residuals_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "SPIN and WhyPlay Adjusted Residuals Heatmap", x = "SPIN Scale", y = "WhyPlay") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
```

With regards to indicators of social phobia, the most apparent indicator is that of "escape" and very severe outcomes. The typical pattern of "having fun" being associated with better outcomes and "winning" having the inverse effect is also present.

```{r SPIN_ord vs why_play}

ggplot(df, aes(x = SPIN_ord, fill = cleaned_whyplay)) +
  geom_bar(position = "fill") +  # Use position = "fill" to show proportions
  ylab("Proportion") +           # Label the y-axis
  xlab("Total score on the Social Phobia Inventory") +          # Label the x-axis
  labs(title = "Proportion of Reasons for Playing by SPIN Scores",
       fill = "Reasons for Playing") +  # Rename the legend title
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),  # Center and bold the title
    axis.title.x = element_text(face = "bold", size = 12),  # Bold and size for x-axis label
    axis.title.y = element_text(face = "bold", size = 12),  # Bold and size for y-axis label
    axis.text.x = element_text(face = "bold", size = 10),   # Bold the x-axis tick labels
    axis.text.y = element_text(face = "bold", size = 10),   # Bold the y-axis tick labels
    legend.title = element_text(face = "bold"),             # Bold the legend title
    legend.text = element_text(face = "bold")               # Bold the legend labels
  )
  

```

### ANOVA test for Total time spent and GAD_ordinal variables

As we have seen before from the histogram distributions before that the hours spent playing games and watching streams is different for different levels of anxiety. Now in this part we will look if the amount of time people spend on gaming and streaming is signifantly different depending on their level of Generalized Anxiety Disorder (GAD). So we state the following hypotheses:

Null hypothesis: There is no difference in the mean combined hours spent on gaming and streaming among the different levels of GAD.
Alternate hypothesis: At least one group mean of the combined hours spent on gaming and streaming is different from the others among the different levels of GAD.


```{r}

# Combine the hours played and streams into a new variable
df$hrs_stream_combined <- df_clean$streams + hrs_GAD_filtered_clean$Hours

# Perform the ANOVA test
anova_result_combined <- aov(hrs_stream_combined ~ GAD_ord, data = df)

# Summarize the ANOVA results
summary(anova_result_combined)

```

The p-value from the test is very small (2.1e-07), which is much smaller than the common threshold of 0.05. Because the p-value is so small, we can confidently say that the difference in time spent gaming and streaming is not just due to random chance. This means that people with different levels of GAD do indeed spend significantly different amounts of time on gaming and streaming.

So, at least one group (for example, those with mild anxiety) spends a different amount of time gaming and streaming compared to another group (for example, those with severe anxiety). Next yo understand which specific groups are different, we want to conduct post-hoc tests (like Tukey's HSD) to compare the means between each pair of GAD levels.

#### Post hoc test

The Tukey HSD test will compare the means of every possible pair of groups (levels of GAD) to see which means are significantly different from each other.


```{r}
# Perform the ANOVA
anova_result <- aov(hrs_stream_combined ~ GAD_ord, data = df)

# Perform Tukey HSD post-hoc test
tukey_result <- TukeyHSD(anova_result)

# Display the Tukey HSD results
print(tukey_result)

# Optional: Plot the Tukey HSD results
plot(tukey_result)

```

The results suggest that individuals with higher levels of GAD (severe and moderate) tend to spend more combined hours on gaming and streaming compared to those with minimal anxiety. However, the difference between some intermediate levels (e.g., moderate vs. mild) is not statistically significant.

## <u>Linear Regression Modeling</u>
We want to make inferences on how mental health indicator scores and categories affect a person's time spent playing and streaming video games. In order to start this process, we are first going to guess that there is a linear relationship among all the variables.

### Numeric Scores
We will first start with a linear model that has individual or combined mental health scores as the independent variable(s) and hours playing and streaming as the dependent variable.

```{r Numeric_LRM}
# Fit the linear regression models
lm_gad <- lm(hrs_stream_combined ~ GAD_T, data = df)
lm_swl <- lm(hrs_stream_combined ~ SWL_T, data = df)
lm_spin <- lm(hrs_stream_combined ~ SPIN_T, data = df)
lm_3_combined <- lm(hrs_stream_combined ~ GAD_T + SWL_T + SPIN_T, data = df)

# Define the summaries of each model
summary_gad <- summary(lm_gad)
summary_swl <- summary(lm_swl)
summary_spin <- summary(lm_spin)
summary_3_combined <- summary(lm_3_combined)

# Values
## Intercepts and slopes
intercept_gad <- coef(summary_gad)["(Intercept)", "Estimate"]
slope_gad <- coef(summary_gad)["GAD_T", "Estimate"]
## P-value for the slope
p_value_gad <- coef(summary_gad)["GAD_T", "Pr(>|t|)"]
## R-squared and Adjusted R-squared
r_squared_gad <- summary_gad$r.squared
adj_r_squared_gad <- summary_gad$adj.r.squared

# Call the summaries
summary_gad
summary_swl
summary_spin
summary_3_combined

# Visualize the relationship between Hours Playing & Streaming and GAD_T
ggplot(df, aes(x = GAD_T, y = hrs_stream_combined)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Linear Regression: Hours Playing and Streaming vs GAD_T",
       x = "Generalized Anxiety Disorder Total Score",
       y = "Hours Spent Playing and Streaming Video Games") +
  theme_minimal()

# Visualize the relationship between Hours Playing & Streaming and SWL_T
ggplot(df, aes(x = SWL_T, y = hrs_stream_combined)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Linear Regression: Hours Playing and Streaming vs SWL_T",
       x = "Satisfaction with Life Total Score",
       y = "Hours Spent Playing and Streaming Video Games") +
  theme_minimal()

# Visualize the relationship between Hours Playing & Streaming and SPIN_T
ggplot(df, aes(x = SPIN_T, y = hrs_stream_combined)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Linear Regression: Hours Playing and Streaming vs SWL_T",
       x = "Social Phobia Inventory Total Score",
       y = "Hours Spent Playing and Streaming Video Games") +
  theme_minimal()

# Visualize the relationship among Hours Playing & Streaming and the three, numeric scores
ggplot(df, aes(x = GAD_T + SWL_T + SPIN_T, y = hrs_stream_combined)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Linear Regression: Hours Playing and Streaming vs GAD_T + SWL_T + SPIN_T",
       x = "3 Total Scores",
       y = "Hours Spent Playing and Streaming Video Games") +
  theme_minimal()
```

#### Outputs of Numeric Models
I will only be talking about the GAD_T model, as similar logic can be applied to SWL_T, SPINT_T, and the combination of the three metrics. The intercept is `r intercept_gad`. The GAD model is telling us that people in our data set that have a total score of 0 for the GAD-7 questionnaire spends approximately `r round(intercept_gad, 0)` hours playing and streaming video games. The slope of the line of best fit is `r slope_gad`, which means that as the score from the questionnaire increases by 1, the amount of hours playing and streaming video games increases by approximately `r round(slope_gad, 1)` hours. However, the scatter plot is showing a huge issue. Players can only receive <b>integer value scores</b> when taking the GAD-7 questionnaire. This makes all other values in the models, like Adjusted R Squared and p-values (`r adj_r_squared_gad` and `r p_value_gad` respectively), useless. Before abandoning linear regression models altogether, let's see how using the ordinal categories for each metric plays out.

### Ordinal Scores
We will see how linear models that have the ordinal categories of each  metric functions as the independent variable(s) and hours playing and streaming as the dependent variable.

```{r Odrinal_LRM}
# Fit the linear regression models
lm_gad_ord <- lm(hrs_stream_combined ~ GAD_ord, data = df)
lm_GADE <- lm(hrs_stream_combined ~ GADE, data = df)
lm_swl_ord <- lm(hrs_stream_combined ~ SWL_ord, data = df)
lm_spin_ord <- lm(hrs_stream_combined ~ SPIN_ord, data = df)
lm_4_ord <- lm(hrs_stream_combined ~ GAD_ord + GADE + SWL_ord + SPIN_ord, data = df)

# Define the summaries of each model
summary_gad_ord <- summary(lm_gad_ord)
summary_GADE <- summary(lm_GADE)
summary_swl_ord <- summary(lm_swl_ord)
summary_spin_ord <- summary(lm_spin_ord)
summary_4_ord <- summary(lm_4_ord)

# Values
## R-squared and Adjusted R-squared
r_squared_gad_ord <- summary_gad_ord$r.squared
adj_r_squared_gad_ord <- summary_gad_ord$adj.r.squared
r_squared_4_ord <- summary_4_ord$r.squared
adj_r_squared_4_ord <- summary_4_ord$adj.r.squared

# Call the summaries
summary_gad_ord
summary_GADE
summary_swl_ord
summary_spin_ord
summary_4_ord

ggplot(df, aes(x = GAD_ord, y = hrs_stream_combined)) +
    geom_boxplot() +
    labs(title = "Boxplot of Total Hours by GAD Levels",
         x = "GAD Ordinal Levels",
         y = "Combined Hours (Gaming + Streaming)") +
    theme_minimal()

ggplot(df, aes(x = GADE, y = hrs_stream_combined)) +
    geom_boxplot() +
    labs(title = "Boxplot of Total Hours by GADE Levels",
         x = "GADE Ordinal Levels",
         y = "Combined Hours (Gaming + Streaming)") +
    theme_minimal()

ggplot(df, aes(x = SWL_ord, y = hrs_stream_combined)) +
    geom_boxplot() +
    labs(title = "Boxplot of Total Hours by SWL Levels",
         x = "SWL Ordinal Levels",
         y = "Combined Hours (Gaming + Streaming)") +
    theme_minimal()

ggplot(df, aes(x = SPIN_ord, y = hrs_stream_combined)) +
    geom_boxplot() +
    labs(title = "Boxplot of Total Hours by SPIN Levels",
         x = "SPIN Ordinal Levels",
         y = "Combined Hours (Gaming + Streaming)") +
    theme_minimal()
```

#### Outputs of Ordinal Models

Similarly with the numeric models, I will mostly only discuss the GAD_T model, as the logic for the other models is similar. Using ordinal categories was much better than raw, numeric scores. We get a clear depiction of median hours spent playing and streaming video games based on the category players align to. However, there is still a significant issue. While everything is set up correctly, we receive an R-squared value of `r r_squared_gad_ord`; the adjusted R-squared value is `r adj_r_squared_gad_ord`. This means that the model only `r round(adj_r_squared_gad_ord, 0)*100`% of the variance in the ordinal GAD categories is accounted for. Even using the model that combines all the ordinal categories, including GADE, yields an adjusted R-squared value of `r adj_r_squared_4_ord`, meaning only `r round(adj_r_squared_4_ord, 0)*100`% of the variance is accounted for. This evidence informs us that there is not a linear relationship among the metrics in the data set and hours spent playing and streaming video games. Linear regression is not the correct type of regression to perform for this analysis, so we will have to use logistic regression.

## <u>Key Findings</u> 

Team 5's analysis reveals significant relationships between video game habits and mental health outcomes. First, increasing hours spent playing video games and streaming is associated with higher stream counts, as indicated by significant ANOVA results across all comparisons. This suggests that gamers who spend more hours playing are more likely to be heavily involved in the gaming community, leading to increased streaming activity.

In examining the impact of anxiety levels on gaming behavior, we found that individuals with Very High anxiety levels engage in significantly more streaming hours compared to those with Low, Moderate, and High anxiety. Conversely, higher levels of gaming hours negatively affect subjective well-being (SWL_T), with significant declines observed at Very High hours played.

Additionally, SPIN_T scores significantly increase from Low to Very High levels of gaming, indicating that greater gaming engagement correlates with heightened psychological distress.

Lastly, our chi-squared analysis shows that respondents who play video games for enjoyment report better mental health outcomes than those motivated by competition or escapism.

As causality cannot be determined, Team 5 caveats this analysis by adding that pre-exiting mental health conditions and other life stressors could be a contributing factor to negative mental health outcomes. However, a correlative effect is clearly seen as exposed by the ANOVA and Chi-Sqaured tests we conducted. 

## <u>Recommendations</U>

Team 5 recommends passing our findings to social scientists to adjust the survey instruments. Collecting more comprehensive demographic information would be invaluable, including factors such as respondents' socioeconomic status (e.g., ability to afford the cost of living), family history of mental health diagnoses, and other relevant contextual data. This information could help identify potential confounding variables and provide deeper insights into the relationships we observed.

Given the non-normality of the inventory scores, we also recommend conducting logistic regression to observe the interaction of multiple predictors against the outcomes (inventory scores). Adding this layer of complexity would allow us to more effectively model how gaming habits affect mental health outcomes in gamers. 

### Tree Based -- Classifcication 

Initial Data prep

``` {r, randomforest_classifcation}
df <- df[complete.cases(df$GAD_ord, df$cleaned_whyplay, df$hrs_stream_combined, df$Game,df$Platform, df$earning, df$Gender, df$Age, df$Work,df$Residence), ]


#Random Forest does not allow chr, they have to be factors
df$cleaned_whyplay <- factor(df$cleaned_whyplay)
df$Game <- factor(df$Game)
df$Gender <- factor(df$Gender)
df$GAD_ord <- factor(as.character(df$GAD_ord))

#encode as needed
cleaned_whyplay_encoded <- as.integer(factor(df$cleaned_whyplay))
gender_encoded <- as.integer(factor(df$Gender))
game_encoded <- as.integer(factor(df$Game))

df$cleaned_whyplay_encoded <- cleaned_whyplay_encoded
df$gender_encoded <- gender_encoded
df$game_encoded <- game_encoded
str(df)
```


### GAD

```{r, smote_GAD}
# Ensure the target variable GAD_ord is a factor
df$GAD_ord <- as.factor(df$GAD_ord)

# Select the relevant predictors and the target variable
predictors <- df[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- df$GAD_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 4)

# Combine the SMOTE result into a new data frame
smote_data <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data)[ncol(smote_data)] <- "GAD_ord"

# Ensure the target variable in the new data frame is a factor
smote_data$GAD_ord <- as.factor(smote_data$GAD_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data$GAD_ord))




```

run smote again for the next minority class (GAD)

```{r, smote_again_gad}
# Ensure the target variable GAD_ord is a factor
df$GAD_ord <- as.factor(df$GAD_ord)

# Select the relevant predictors and the target variable
predictors <- smote_data[, c("cleaned_whyplay_encoded", "Hours", "streams", "gender_encoded", "game_encoded", "Age")]
target <- smote_data$GAD_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 2)

# Combine the SMOTE result into a new data frame
smote_data2 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data2)[ncol(smote_data)] <- "GAD_ord"

# Ensure the target variable in the new data frame is a factor
smote_data2$GAD_ord <- as.factor(smote_data2$GAD_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data2$GAD_ord))


class_counts <- table(smote_data2$GAD_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of GAD_ord Classes",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))

str(smote_data2)
```


```{r, model_gad}
#model_bal <- randomForest(GAD_ord ~ cleaned_whyplay_encoded + Hours * streams #,
 #                            data = smote_data2, 
 #                            ntree = 1000)

#cat("Accuracy -- bal: ", mean(predict(model_bal) == smote_data2$GAD_ord), #"\n")

#print(model_bal)

#importance(model_bal)
#varImpPlot(model_bal)
```

now undersample the majority class

```{r, undersample_gad}
# Check the distribution of the classes
table(smote_data2$GAD_ord)

# Ensure the "GAD_ord" column is a factor or character
smote_data2$GAD_ord <- as.character(smote_data2$GAD_ord)  # Convert to character if it's a factor

# Now, under-sample the "minimal anxiety" class to 3643
set.seed(123)  # For reproducibility

# Subset the "minimal anxiety" rows and the other classes
minimal_anxiety <- smote_data2[smote_data2$GAD_ord == "minimal anxiety", ]
other_classes <- smote_data2[smote_data2$GAD_ord != "minimal anxiety", ]

# Ensure there are enough rows to sample from the "minimal anxiety" class
if (nrow(minimal_anxiety) >= 3643) {
  minimal_anxiety_undersampled <- minimal_anxiety[sample(1:nrow(minimal_anxiety), 3643), ]
} else {
  stop("Not enough rows in 'minimal anxiety' class to sample 3643.")
}

# Combine the undersampled "minimal anxiety" class with the other classes
df_undersampled <- rbind(minimal_anxiety_undersampled, other_classes)

# Check the new distribution
table(df_undersampled$GAD_ord)


class_counts <- table(df_undersampled$GAD_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of GAD_ord Classes",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))


df_undersampled$GAD_ord <- as.factor(df_undersampled$GAD_ord)
```

run the model again

```{r, balmodel}
# Consolidate minimal and mild anxiety into "minimal-mild anxiety"
df_undersampled$GAD_ord <- factor(
  ifelse(df_undersampled$GAD_ord %in% c("minimal anxiety", "mild anxiety"), 
         "minimal-mild anxiety", 
         as.character(df_undersampled$GAD_ord)), # Retain other classes
  levels = c("minimal-mild anxiety", "moderate anxiety", "severe anxiety") # Define desired order
)

# Verify structure
str(df_undersampled)

#levels(df_undersampled$GAD_ord)


model_bal2 <- randomForest(GAD_ord ~ cleaned_whyplay_encoded + Hours + streams + game_encoded + gender_encoded + Age ,
                             data = df_undersampled, 
                             ntree = 3000,
                            mtry = 3, 
                            nodesize = 5, 
                         importance = TRUE)



cat("Accuracy -- bal: ", mean(predict(model_bal2) == df_undersampled$GAD_ord), "\n")

print(model_bal2)

#importance(model_bal2)
varImpPlot(model_bal2)

minimal_mild <- df[df$GAD_ord %in% c("minimal anxiety", "mild anxiety"), ]

# Summary statistics of GAD_T for each category
aggregate(GAD_T ~ GAD_ord, data = minimal_mild, summary)



# Violin plot to visualize the distribution with y-axis scaled from 0 to 9
ggplot(minimal_mild, aes(x = GAD_ord, y = GAD_T, fill = GAD_ord)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Violin Plot of GAD_T for Minimal and Mild Anxiety",
       x = "Anxiety Category", y = "GAD_T Score", fill = "Anxiety Category") +
  ylim(0, 9) +
  theme_minimal()





```

### SWL

```{r, smote_swl}
# Ensure the target variable GAD_ord is a factor
df$SWL_ord <- as.factor(df$SWL_ord)
print(table(df$SWL_ord))

class_counts <- table(df$SWL_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of SWL_ord Classes",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))


# Select the relevant predictors and the target variable
predictors <- df[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- df$SWL_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 4)

# Combine the SMOTE result into a new data frame
smote_data <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data)[ncol(smote_data)] <- "SWL_ord"

# Ensure the target variable in the new data frame is a factor
smote_data$SWL_ord <- as.factor(smote_data$SWL_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data$SWL_ord))

class_counts <- table(smote_data$SWL_ord)



class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SWL_ord Classes, 1st pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))



```

run smote again for the next minority class (SWL)

```{r, smote_again_SWL}



# Select the relevant predictors and the target variable
predictors <- smote_data[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- smote_data$SWL_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 4)

# Combine the SMOTE result into a new data frame
smote_data2 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data2)[ncol(smote_data2)] <- "SWL_ord"

# Ensure the target variable in the new data frame is a factor
smote_data2$SWL_ord <- as.factor(smote_data2$SWL_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data2$SWL_ord))

class_counts <- table(smote_data2$SWL_ord)



class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SWL_ord Classes, 2nd pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))




```

run smote for a third time to ge the last minority class

```{r, smotex3_swl}

# Select the relevant predictors and the target variable
predictors <- smote_data2[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- smote_data2$SWL_ord

cat("Rows in predictors:", nrow(predictors), "\n")
cat("Rows in target:", length(target), "\n")

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 2)

# Combine the SMOTE result into a new data frame
smote_data3 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data3)[ncol(smote_data3)] <- "SWL_ord"

# Ensure the target variable in the new data frame is a factor
smote_data3$SWL_ord <- as.factor(smote_data3$SWL_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data3$SWL_ord))

class_counts <- table(smote_data3$SWL_ord)



class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SWL_ord Classes, 3rd pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))




```

```{r, undersmaple_swl}


# Ensure the "GAD_ord" column is a factor or character
smote_data3$SWL_ord <- as.character(smote_data3$SWL_ord)  # Convert to character if it's a factor

# Now, under-sample the "minimal anxiety" class to 3643
set.seed(123)  # For reproducibility

# Subset the "minimal anxiety" rows and the other classes
extremely_satisfied <- smote_data3[smote_data3$SWL_ord == "extremely satisfied", ]
other_classes <- smote_data3[smote_data3$SWL_ord != "extremely satisfied", ]


if (nrow(extremely_satisfied) >= 2851) {
  extreme_sat_undersampled <- extremely_satisfied[sample(1:nrow(extremely_satisfied), 2851), ]
} else {
  stop("Not enough rows in 'minimal anxiety' class to sample 2851.")
}

# Combine the undersampled "minimal anxiety" class with the other classes


df_undersampled_swl <- rbind(extreme_sat_undersampled, other_classes)

# Check the new distribution
table(df_undersampled_swl$SWL_ord)


class_counts <- table(df_undersampled_swl$SWL_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of SWL_ord, after undersamp",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))


#df_undersampled_swl$SWL_ord <- as.factor(df_undersampled$SWL_ord)
```

```{r, more_robust_us}
balance_classes <- function(df_undersampled_swl, target_column = "SWL_ord", reference_class = "dissatisfied") {
  # Ensure the target column exists in the dataframe
  if (!target_column %in% colnames(df_undersampled_swl)) {
    stop(paste("Column", target_column, "does not exist in the dataframe"))
  }

  # Identify the size of the reference class (dissatisfied)
  dissatisfied_size <- sum(df_undersampled_swl[[target_column]] == reference_class)
  
  # List to store the undersampled data
  undersampled_data <- list()

  # Loop over each class and sample from it to match the dissatisfied class size
  for (class_label in unique(df_undersampled_swl[[target_column]])) {
    # Subset data for the current class
    class_data <- df_undersampled_swl[df_undersampled_swl[[target_column]] == class_label, ]
    
    # If the class size is greater than the dissatisfied class size, sample to match it
    if (nrow(class_data) > dissatisfied_size) {
      class_data <- class_data[sample(1:nrow(class_data), dissatisfied_size), ]
    }
    
    # Append the undersampled data
    undersampled_data[[class_label]] <- class_data
  }
  
  # Combine all undersampled data into a single dataframe
  balanced_data <- do.call(rbind, undersampled_data)
  
  return(balanced_data)
}

balanced_data <- balance_classes(df_undersampled_swl, target_column = "SWL_ord", reference_class = "dissatisfied")

# Check the structure and distribution of the resulting dataset
str(balanced_data)
table(balanced_data$SWL_ord)

```

```{r, violin_swl}

minimal_mild <- df[df$SWL_ord %in% c("extremely satisfied", "satisfied", "slightly satisfied", "neutral", "slightly dissatisfied", "dissatisfied", "extremely dissatisfied" ), ]

# Summary statistics of GAD_T for each category
aggregate(SWL_T ~ SWL_ord, data = minimal_mild, summary)

# Violin plot to visualize the distribution with y-axis scaled from 0 to 9
ggplot(minimal_mild, aes(x = SWL_ord, y = SWL_T, fill = SWL_ord)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Violin Plot of SWL_T for all classes",
       x = "SWL Category", y = "SWL_T Score", fill = "SWL Category") +

  theme_minimal()
```

```{r, SWL_consolidate}
# Consolidate satisfaction and dissatisfaction levels
df_undersampled_swl$SWL_ord <- factor(
  ifelse(df_undersampled_swl$SWL_ord %in% c("satisfied", "slightly satisfied"), 
         "sat-slightly_sat", 
         ifelse(df_undersampled_swl$SWL_ord %in% c("dissatisfied", "slightly dissatisfied"), 
                "dis-slightly_dis", 
                as.character(df_undersampled_swl$SWL_ord))), # Retain other classes
  levels = c("extremely satisfied", "sat-slightly_sat", "neutral", 
             "dis-slightly_dis", "extremely dissatisfied") # Define desired order
)

# Verify structure
str(df_undersampled_swl)

# Check the levels
levels(df_undersampled_swl$SWL_ord)

table(df_undersampled_swl$SWL_ord)



ggplot(df_undersampled_swl, aes(x = Hours, fill = SWL_ord)) +
  geom_density(alpha = 0.5) +
  labs(title = "Feature Distribution by Class", x = "Hours", y = "Density")
```

```{r, SWL_model}



df_undersampled_swl$SWL_ord <- as.factor(df_undersampled_swl$SWL_ord)

model_bal_swl <- randomForest(SWL_ord ~ cleaned_whyplay_encoded + Hours + streams + game_encoded + gender_encoded + Age ,
                             data = df_undersampled_swl, 
                            ntree = 3000,
                           mtry = 4, 
                        nodesize = 5, 
                         importance = TRUE)



cat("Accuracy -- bal: ", mean(predict(model_bal_swl) == df_undersampled_swl$SWL_ord), "\n")

print(model_bal_swl)

#importance(model_bal2)
varImpPlot(model_bal_swl)




```

****************
```{r, swl_hyperparams}
# Assuming `balanced_data` is your dataset
#set.seed(123)  # For reproducibility
#trainIndex <- createDataPartition(df_undersampled_swl$SWL_ord, p = 0.8, list = FALSE)
#train_data <- df_undersampled_swl[trainIndex, ]
#test_data <- df_undersampled_swl[-trainIndex, ]
#test_data$SWL_ord <- as.factor(test_data$SWL_ord)

#grid <- expand.grid(
#  mtry = c(2, 4, 6),             # Number of variables to try at each split
#  splitrule = "gini",            # Splitting criterion (if using `ranger`)
#  min.node.size = c(3, 5, 7, 9)     # Minimum node size
#)

#control <- trainControl(
#  method = "cv",          # Cross-validation
#  number = 5,             # Number of folds
#  verboseIter = TRUE      # Display training progress
#)


# Train the Random Forest model
#rf_model <- train(
#  SWL_ord ~ cleaned_whyplay_encoded + Hours + streams + game_encoded + gender_encoded + Age,
#  data = train_data,
#  method = "ranger",         # Using ranger implementation for efficiency
#  trControl = control,
#  tuneGrid = grid
#)

# Display the best parameters and results
#print(rf_model)


# Predictions on the test dataset
#predictions <- predict(rf_model, newdata = test_data)


```

*************





### SPIN

```{r, violin_spin}


minimal_mild <- df[df$SPIN_ord %in% c("none", "mild", "moderate", "severe", "very severe"), ]

# Summary statistics of GAD_T for each category
aggregate(SPIN_T ~ SPIN_ord, data = minimal_mild, summary)

# Violin plot to visualize the distribution with y-axis scaled from 0 to 9
ggplot(minimal_mild, aes(x = SPIN_ord, y = SPIN_T, fill = SPIN_ord)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Violin Plot of SPNI_T for all classes",
              x = "SPIN Category", y = "SPIN_T Score", fill = "SPIN Category") +
  theme_minimal()
```

```{r, smote1_spin}


df$SPIN_ord <- as.factor(df$SPIN_ord)
print(table(df$SPIN_ord))

class_counts <- table(df$SPIN_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of SPIN_ord Classes, Initial",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))


# Select the relevant predictors and the target variable
#predictors <- df[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
#target <- df$SPIN_ord

#cat("Rows in predictors:", nrow(predictors), "\n")
#cat("Rows in target:", length(target), "\n")

#str(predictors)
#str(target)

# Check for any NA values or data inconsistencies
#sum(is.na(predictors))
#sum(is.na(target))

df <- df[!is.na(df$SPIN_ord), ]

#target <- df$SPIN_ord

sum(is.na(target))

# Apply SMOTE
#set.seed(123) # For reproducibility
#smote_result2 <- SMOTE(X = predictors, target = target, K = 2, dup_size = 1)

# Ensure the target variable GAD_ord is a factor
df$SPIN_ord <- as.factor(df$SPIN_ord)

# Select the relevant predictors and the target variable
predictors <- df[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- df$SPIN_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 4)



# Combine the SMOTE result into a new data frame
smote_data_spin1 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data_spin1)[ncol(smote_data_spin1)] <- "SPIN_ord"

# Ensure the target variable in the new data frame is a factor
smote_data_spin1$SWL_ord <- as.factor(smote_data_spin1$SPIN_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data_spin1$SPIN_ord))

str(smote_data_spin1)

class_counts <- table(smote_data_spin1$SPIN_ord)

print(class_counts)

class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SPIN_ord Classes, 1st pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))

```

```{r, spin_smote2}
# Ensure the target variable GAD_ord is a factor
smote_data_spin1$SPIN_ord <- as.factor(smote_data_spin1$SPIN_ord)

# Select the relevant predictors and the target variable
predictors <- smote_data_spin1[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- smote_data_spin1$SPIN_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 2)



# Combine the SMOTE result into a new data frame
smote_data_spin2 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data_spin2)[ncol(smote_data_spin2)] <- "SPIN_ord"

# Ensure the target variable in the new data frame is a factor
smote_data_spin2$SWL_ord <- as.factor(smote_data_spin2$SPIN_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data_spin2$SPIN_ord))

str(smote_data_spin2)

class_counts <- table(smote_data_spin2$SPIN_ord)

print(class_counts)

class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SPIN_ord Classes, 2nd pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))

```

```{r, lastsmote_mod}
# Ensure the target variable GAD_ord is a factor
smote_data_spin2$SPIN_ord <- as.factor(smote_data_spin2$SPIN_ord)

# Select the relevant predictors and the target variable
predictors <- smote_data_spin2[, c("cleaned_whyplay_encoded", "Hours", "streams", "game_encoded", "gender_encoded", "Age")]
target <- smote_data_spin2$SPIN_ord

# Apply SMOTE
set.seed(123) # For reproducibility
smote_result <- SMOTE(X = predictors, target = target, K = 1, dup_size = 1)



# Combine the SMOTE result into a new data frame
smote_data_spin3 <- data.frame(smote_result$data)

# Rename the target variable
names(smote_data_spin3)[ncol(smote_data_spin3)] <- "SPIN_ord"

# Ensure the target variable in the new data frame is a factor
smote_data_spin3$SWL_ord <- as.factor(smote_data_spin3$SPIN_ord)

# Verify the distribution of the target variable after SMOTE
print(table(smote_data_spin3$SPIN_ord))

str(smote_data_spin3)

class_counts <- table(smote_data_spin3$SPIN_ord)

print(class_counts)

class_percentages <- prop.table(class_counts) * 100


# Pie chart of class percentages
pie(class_counts,
   main = "Percentage of SPIN_ord Classes, 3rd pass",
    col = rainbow(length(class_counts)),
  labels = paste(names(class_counts), round(class_percentages, 1), "%"))
```

```{r, spin_under}

# Ensure the "GAD_ord" column is a factor or character
smote_data_spin3$SPIN_ord <- as.character(smote_data_spin3$SPIN_ord)  # Convert to character if it's a factor

# Now, under-sample the "minimal anxiety" class to 3643
set.seed(123)  # For reproducibility

# Subset the "minimal anxiety" rows and the other classes
extremely_satisfied <- smote_data_spin3[smote_data_spin3$SPIN_ord == "none", ]
other_classes <- smote_data_spin3[smote_data_spin3$SPIN_ord != "none", ]


if (nrow(extremely_satisfied) >= 3500) {
  extreme_sat_undersampled <- extremely_satisfied[sample(1:nrow(extremely_satisfied), 3500), ]
} else {
  stop("Not enough rows in 'minimal anxiety' class to sample 2851.")
}

# Combine the undersampled "minimal anxiety" class with the other classes


df_undersampled_spin <- rbind(extreme_sat_undersampled, other_classes)

# Check the new distribution
table(df_undersampled_spin$SPIN_ord)


class_counts <- table(df_undersampled_spin$SPIN_ord)

class_percentages <- prop.table(class_counts) * 100

# Pie chart of class percentages
pie(class_counts,
    main = "Percentage of SPIN_ord, after undersamp",
    col = rainbow(length(class_counts)),
    labels = paste(names(class_counts), round(class_percentages, 1), "%"))


```



```{r, spin_consolidate}
# Consolidate 'mild' and 'moderate' into 'mild-moderate'
df_undersampled_spin$SPIN_ord <- factor(
  ifelse(df_undersampled_spin$SPIN_ord %in% c("mild", "moderate"), 
         "mild-moderate", 
         as.character(df_undersampled_spin$SPIN_ord)),  # Retain other classes
  levels = c("none", "mild-moderate", "severe", "very severe")  # Define desired order
)

# Verify structure
str(df_undersampled_spin)

# Check the levels
levels(df_undersampled_spin$SPIN_ord)

# Check the distribution of the consolidated class
table(df_undersampled_spin$SPIN_ord)

# Visualize the feature distribution by the new consolidated class
ggplot(df_undersampled_spin, aes(x = Hours, fill = SPIN_ord)) +
  geom_density(alpha = 0.5) +
  labs(title = "Feature Distribution by Class (Consolidated)", x = "Hours", y = "Density")

```


```{r, spinmodel}
df_undersampled_spin$SPIN_ord <- as.factor(df_undersampled_spin$SPIN_ord)


#class_weights <- c("none" = 3, "mild" = 3, "moderate" = 2, "severe" = 1, "very severe" = 1)
model_bal_spin <- randomForest(SPIN_ord ~ cleaned_whyplay_encoded + Hours + streams + game_encoded + gender_encoded + Age ,
                             data = df_undersampled_spin, 
                            ntree = 3000,
                           mtry = 4, 
                        nodesize = 3, 
                         importance = TRUE
                       )



cat("Accuracy -- bal: ", mean(predict(model_bal_spin) == df_undersampled_spin$SPIN_ord), "\n")

print(model_bal_spin)

#importance(model_bal2)
varImpPlot(model_bal_spin)
```




## <u>References</u>



## <u>References</u>










